---
title: "Chapter 3 - The Nature of Solar Flare Occurrences"
author: "Joshua Ingram"
date: "5/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(wesanderson)
library(car)
library(VGAM)
library(qqplotr)
library(lubridate)
library(zoo)
library(goftest)
library(latex2exp)
library(padr)
source('D:/main/Projects/Solar-Flare-Project/R/exp_sim_ks_stat.R')

GOES <- read_csv("D:/main/Datasets/Solar_REU/GOES_data/goes_clean.csv")
GOES$cycle <- as.factor(GOES$cycle)
```

# 3.1 Solar Flare Occurrences as a Poisson Process

Barplot of flare frequences across solar cycles 23 and 24
```{r}
# preparing histograms for flare frequencies
dates <- c(as.Date("1996-07-30 00:00:01 UTC"), 
           as.Date("2008-12-01 00:00:01 UTC"), 
           as.Date("2001-11-15 00:00:01 UTC"), 
           as.Date("2014-04-15 00:00:01 UTC"))
extrema <- c("Minimum", "Minimum", "Maximum", "Maximum")
solar_extrema <- data.frame(dates, extrema)

ggplot(data = GOES, aes(x=year, fill = as.factor(cycle))) + 
  geom_bar(color = "black") +
  geom_vline(data = solar_extrema, mapping = aes(xintercept = decimal_date(dates), linetype = extrema), size = .85) +
  labs(x = "Year", y = "Count", fill = "Cycle", linetype = "Extrema") +
  theme_bw() +
  scale_fill_brewer(palette = "Set2")
```

Barplot of flare frequenies by class across solar cycles 23 and 24
```{r}
ggplot(data = GOES %>% mutate(class= recode(class, "c('A', 'B')='A-B'")), aes(x=year, fill = as.factor(cycle))) + 
  geom_bar(color = "black") +
  geom_vline(data = solar_extrema, mapping = aes(xintercept = decimal_date(dates), linetype = extrema), size = .85) +
  facet_wrap(~class) +
  scale_y_log10() +
  labs(x = "Year", y = "Count", fill = "Cycle", linetype = "Extrema") +
  theme_bw() +
  scale_fill_brewer(palette = "Set2")
```


# 3.2 Evaluating Goodness of Fits

```{r}
# random sample of sample size 100 from exponential(5)
set.seed(10)
exp_sample <- rexp(100, rate = 5)
sim_data <- data.frame(sample = exp_sample)
lambda <- 1/mean(exp_sample)

ggplot(sim_data, aes(sample = sample)) +
  stat_pp_band(distribution = "exp") +
  stat_pp_line(color = "sienna2") +
  stat_pp_point(distribution = "exp",
                color = "aquamarine4") +
  theme_bw() +
  labs(title = "P-P Plot for Exponential Fit to Sample Drawn from Exponential", subtitle = "This consititutes a good fit", y = "Empirical Cumulative Probability", x = "Theoretical Cumulative Probability")

set.seed(10)
log_sample <- rlnorm(100, meanlog = 0, sdlog = 2)
sim_data <- data.frame(sample = log_sample)
lambda <- 1/mean(log_sample)

ggplot(sim_data, aes(sample = sample)) + 
  stat_pp_band(distribution = "exp") +
  stat_pp_line(color = "sienna2") +
  stat_pp_point(distribution = "exp",
                color = "aquamarine4") +
  theme_bw() +
  labs(title = "P-P Plot for Exponential Fit to Sample Drawn from Lognormal", subtitle = "This does not constitute a good fit: overdispersed", y = "Empirical Cumulative Probability", x = "Theoretical Cumulative Probability")
```


# 3.3 Distribution of Flare Counts By Year

Monthly Flare Counts
```{r}
# monthly flare counts
GOES_counts <- GOES %>% mutate(yearmonth = as.Date(as.yearmon(format(Gstart, format = "%Y-%m")))) %>% count(yearmonth, cycle)
GOES_counts <- rbind(c("1996-06-01", 23, 0), GOES_counts) %>% pad %>% fill_by_value(yearmonth) %>% replace_na(list(n = 0)) %>% filter(yearmonth > "1996-06-01")
GOES_counts$n <- as.numeric(GOES_counts$n)
GOES_counts$cycle[GOES_counts$yearmonth < as.Date(as.yearmon("2008-12-01"))] <- 23
GOES_counts$cycle[GOES_counts$yearmonth > as.Date(as.yearmon("2008-12-01"))] <- 24

ggplot(data = GOES_counts, aes(x = yearmonth, y = n, color = cycle)) + 
  geom_vline(data = solar_extrema, mapping = aes(xintercept = dates, linetype = extrema),  size = 0.75) +
  geom_point(size = 1.6) +
  theme_bw() + 
  labs(title = "Monthly Flare Counts", x = "Year", y = "Count", color = "Cycle") +
  scale_color_brewer(palette = "Set2")
```

```{r}
# creating a new variable for the observed flares energy_bin
GOES_bin <- GOES %>% filter(year > 1996) %>% select(Gflrtotalenergy, year, cycle, Gstart)

# filtered out 16 flares that occurred on February 29 (leap year days)
GOES_daily <- GOES_bin %>% 
  mutate(day = day(Gstart), month = month(Gstart)) %>% filter(!(day == 29 & month == 2)) %>% 
  select(-day, -month)

# created variable day that gives the calendar day of each flare
GOES_daily <- GOES_daily %>% 
  mutate(day = yday(Gstart)) %>%
  select(-Gstart)

leap_years <- seq(1996, 2020, 4)

break_levels <- levels(cut(seq(1,370,10), breaks = seq(0, 370, 10)))

# for years in leap_years, for days >= 60, subtract 1
# to look at by year counts, do group_by(year)
GOES_daily <- GOES_daily %>% 
  mutate(day = if_else(year %in% leap_years & day >=60, day - 1, day)) %>% 
  mutate(day_bin = cut(day, breaks = seq(0, 370, 10))) %>% mutate(day_bin = factor(day_bin, levels = break_levels)) %>% complete(day_bin)

GOES_year_count <- GOES_daily %>% group_by(year) %>% count(day_bin, name = "n", .drop = F)
```

```{r}
ggplot(data = GOES_year_count, aes(x = day_bin, y = n)) + 
  geom_point(color = "sienna2") +
  facet_wrap(vars(year)) + 
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  labs(x = "Day Bin", y = "Count")
```

```{r}
# finding mean and variance for each bin range 1999 to 2002
means <- GOES_year_count %>% group_by(year) %>% summarise(mean = mean(n),var = var(n), lower = qpois(.025, lambda = mean), upper = qpois(.975, lambda = mean))

ggplot(data = GOES_year_count, aes(x = day_bin, y = n)) + 
  geom_point( color = "sienna2") + 
  guides(color = FALSE) +
  facet_wrap(vars(year)) + 
  theme_bw() + 
  labs(title = "10-Day Counts for Each Year with Poisson Fits", subtitle = "Solid line is the mean, horizong lines are 2.5 and 97.5 percentiles of fitted Poisson distribution", x = "Day Bin", y = "Count", color = "Energy Bin") +
  geom_hline(data = means, aes(yintercept = mean)) +
  geom_hline(data = means, aes(yintercept = lower), linetype = "dashed") +
  geom_hline(data = means, aes(yintercept = upper), linetype = "dashed") +
  geom_rug(sides = "l") +
  theme(axis.text.x = element_blank())
```

```{r}
ggplot(GOES_year_count, aes(sample = n)) + 
  facet_wrap(vars(year), scales = "free") +
  stat_pp_point(distribution = "pois",
                color = "aquamarine4") +
  stat_pp_line(color = "sienna2") +
  stat_pp_band(distribution = "pois") +
  theme_bw() +
  labs(title = "P-P Plots for Poisson Fits to Yearly 10-Day Count Distributions", x = "Theoretical Cumulative Probability" , 
       y = "Empirical Cumulative Probability")
```

```{r}
# Simulation-based KS test (for exponential only)... Also known as the Lilliefors Test
ptest.sim <- function(data, n = 10000){
  
  # lf - lilliefors, lf_boot - lilliefors double bootstrap, ad_sim - Anderson-Darling like lilliefors, ad_boot - ad double boot
  
  # mle parameter estimate
  lambda <- mean(data)
  
  # ks p-value from original data
  ks_result <- ks.test(data,"ppois",lambda)$p.value
  ad_result <- ad.test(data, "ppois", lambda)$p.value
    
  # simule data from ~exp(lambda) n times, then find p-value
  sim_values <- replicate(n,
                          {
                            sim_data <- rpois(length(data), lambda = lambda)
                            p_val_ks <- ks.test(sim_data, "ppois", lambda = mean(sim_data))$p.value
                            p_val_ad <- ad.test(sim_data, "ppois", lambda = mean(sim_data))$p.value
                            c(p_val_ks, p_val_ad)
                          })
    
  
  sim_values_ks <- c(sim_values[1,], ks_result)
  sim_values_ad <- c(sim_values[2,], ad_result)
  
  # proportion of p-values <= original ks pval... corrected p-value to report
  sim_pval_ks <- mean(sim_values_ks < ks_result)
  sim_pval_ad <- mean(sim_values_ad < ad_result)
  
  
  return(c(sim_pval_ks, sim_pval_ad))
  
}
```


```{r message=FALSE, warning=FALSE}
means <- GOES_year_count %>% group_by(year) %>% summarise(mean = mean(n),var = var(n), inflation_factor = var(n)/mean(n), lower = qpois(.025, lambda = mean), upper = qpois(.975, lambda = mean))

df_results <- data.frame(year = integer(0), ks_pval = integer(0), ks_reject = integer(0), ad_pval = integer(0), ad_reject = integer(0))

# conducting Lilliefors and AD-Lilliefors test for each year
for (i in 1:nrow(means)){
  
  year <- means[i,]$year
  data <- GOES_year_count[which(GOES_year_count$year == year),]$n
  
  result <- ptest.sim(data, n = 10000)
  # Lilliefors version of KS test
  ks_result <- result[1]
  # A Lilliefors-like method for the Anderson-Darling Test 
  ad_result <- result[2]
  
  reject_ks <- ks_result < 0.05
  reject_ad <- ad_result < 0.05
  
  df_results[i,] <- c(year, ks_result, reject_ks, ad_result, reject_ad)
  
}

df_results
mean(df_results$ks_reject)
mean(df_results$ad_reject)
```

# 3.4 Solar Flare Occurrences within Active Regions

Monthly active region counts
```{r}
GOES_AR <- GOES %>% filter(Garreg > 0)
GOES_select <- GOES_AR %>% select(Garreg, year, cycle, Gstart, Glat) %>% group_by(Garreg) %>% filter(row_number() == 1) %>% ungroup()

GOES_ar_count <- GOES_select %>% mutate(yearmonth = as.Date(as.yearmon(format(Gstart, format = "%Y-%m")))) %>% count(yearmonth, cycle)
GOES_ar_count <- rbind(c("1996-06-01", 23, 0), GOES_ar_count) %>% pad %>% fill_by_value(yearmonth) %>% replace_na(list(n = 0)) %>% filter(yearmonth > "1996-06-01")
GOES_ar_count$n <- as.numeric(GOES_ar_count$n)
GOES_ar_count$cycle[GOES_ar_count$yearmonth < as.Date(as.yearmon("2008-12-01"))] <- 23
GOES_ar_count$cycle[GOES_ar_count$yearmonth > as.Date(as.yearmon("2008-12-01"))] <- 24

# Does not include Active Region 0
ggplot(data = GOES_ar_count, aes(x = yearmonth, y = n, color = cycle)) + 
  geom_point() +
  theme_bw() + 
  labs(x = "Year", y = "Count") +
  scale_color_brewer(palette = "Set2")
```

Active region butterfly plot
```{r}
GOES_select <- GOES_AR %>% select(Garreg, year, cycle, Gstart, Glat) %>% group_by(Garreg) %>% filter(row_number() == 1) %>% ungroup()
# removed all ARs with mean lattitude of 0
GOES_ar_lat <- GOES_select %>% 
  mutate(yearmonth = as.Date(as.yearmon(format(Gstart, format = "%Y-%m")))) %>%
  group_by(Garreg) %>% 
  mutate(mean_lat = mean(Glat)) %>% 
  filter(mean_lat != 0)

ggplot(GOES_ar_lat, aes(x = yearmonth, y = mean_lat, color = cycle)) + 
  geom_point() +
  theme_bw() + 
  labs(x = "Year", y = "Average Latitude") +
  scale_color_brewer(palette = "Set2")
```


distribution of number of flares occurring within active regions (percent flares with no AR)
```{r}
GOES_AR <- GOES %>% filter(Garreg > 0)
GOES_select <- GOES_AR %>% select(Garreg, year, cycle, Gstart, Glat) %>% group_by(Garreg) %>% filter(row_number() == 1) %>% ungroup()
GOES_ar_flare_count <- GOES_AR %>% select(Garreg, year, cycle, Gstart, Glat) %>% count(Garreg)

ggplot(data = GOES_ar_flare_count, aes(x = n)) +
  geom_histogram(fill = "aquamarine4", color = "black", binwidth = 3, boundary = 0) +
  theme_bw() + 
  labs(title = "Distribution of Flare Counts within Active Regions", x = "Number of Flares Within Active Region", y = "Count")
```


cases of distribution of waiting times within active regions (make sure to find waiting times between flares within ARs)
```{r}
GOES_inters <- GOES %>% select(Gevtnum, Garreg, year, Gstart, Gstop) %>% group_by(Garreg) %>% arrange(Gstart) %>% mutate(interarrival_time = as.numeric(difftime(Gstart, lag(Gstart), units = "hours"))) %>% drop_na() %>% filter(interarrival_time > 0)

# top ARs
ar_nums <- GOES_ar_flare_count %>% filter(n > 40 & n < 110)
ar_nums <- ar_nums$Garreg[51:70]

ggplot(GOES_inters %>% filter(Garreg %in% ar_nums), aes(x = interarrival_time)) + 
  facet_wrap(vars(Garreg), ncol = 4) +
  geom_histogram(fill = "aquamarine4", color = "black", binwidth = 3, boundary = 0) + 
  theme_bw() + 
  labs(x = TeX("Waiting Time"), y = "Count")
```


p-p plots for n > 79 (only 12 p-p plots) for exponential
```{r}
# top ARs
ar_nums <- GOES_ar_flare_count %>% filter(n > 60)
ar_nums <- ar_nums$Garreg[1:20]

ggplot(GOES_inters %>% filter(Garreg %in% ar_nums), aes(sample = interarrival_time)) + 
  facet_wrap(vars(Garreg), scales = "free", ncol = 4) +
  stat_pp_point(distribution = "exp",
                color = "aquamarine4") +
  stat_pp_line(color = "sienna2") +
  stat_pp_band(distribution = "exp") +
  theme_bw() +
  labs(x = "Theoretical Cumulative Probability" , 
       y = "Empirical Cumulative Probability")
```

p-p plots for n > 79 (only 12 p-p plots) for exponential
```{r}
ar_nums <- df.30$ar_num[20:40]
ar_test <- df.30[c(1,7)] %>% filter(ar_num %in% ar_nums)
ar_test_true <- ar_test %>% filter(ad_exp_true == 1)
ar_test_true <- ar_test_true[,1]
ar_info <- GOES_inters %>% filter(Garreg %in% ar_nums) %>% mutate(exp_true = ifelse(Garreg %in% ar_test_true, 1, 0))

ggplot(ar_info, aes(x = interarrival_time, fill = as.factor(exp_true))) + 
  facet_wrap(vars(Garreg), ncol = 5) +
  geom_histogram(color = "black", binwidth = 3, boundary = 0) + 
  theme_bw() + 
  labs(x = TeX("Interarrival Time"), y = "Count")

# top ARs
ar_nums <- GOES_ar_flare_count %>% filter(n > 30 & n < 35)
ar_nums <- ar_nums$Garreg[1:20]

ggplot(ar_info %>% filter(Garreg %in% ar_nums), aes(sample = interarrival_time, fill = as.factor(exp_true))) + 
  facet_wrap(vars(Garreg), scales = "free", ncol = 4) +
  stat_pp_band(distribution = "exp", alpha = 0.25) +
  stat_pp_line(color = "sienna2") +
  stat_pp_point(distribution = "exp",
                color = "aquamarine4") +
  theme_bw() +
  labs(x = "Theoretical Cumulative Probability" , 
       y = "Empirical Cumulative Probability",
       fill = "Exponential") + 
  scale_fill_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))
```


table of ks, ad tests and percent retained, rejected for AR counts > 20, 30, 50, percent retained
```{r message=TRUE, warning=TRUE}
GOES_inters <- GOES %>% select(Garreg, year, Gpeak, Gstart, Gstop, Glat, Gflrtotalenergy) %>% group_by(Garreg) %>% arrange(Gstart) %>% mutate(interarrival_time = as.numeric(difftime(Gstart, lag(Gstart), units = "hours"))) %>% drop_na() %>% filter(interarrival_time > 0)

df.30 <- compData <- data.frame(ar_num = numeric(0), n = numeric(0), lambda = numeric(0), ks_exp_true = numeric(0), ks_stat = numeric(0), ks_pval = numeric(0), ad_exp_true = numeric(0), ad_stat = numeric(0), ad_pval = numeric(0))
for (i in 1:length(unique(GOES_inters$Garreg))){
  
  ar.num <- unique(GOES_inters$Garreg)[i]
  
  ar_data <- GOES_inters %>% filter(Garreg == ar.num)
  # ar_data$interarrival_time <- ar_data$interarrival_time + rnorm(1, 0, 0.0001)
  n <- nrow(ar_data)
  if (n >= 30){
    
    lambda <- 1/mean(ar_data$interarrival_time)
    results <- gtest.sim(ar_data$interarrival_time, n = 10000, boot = FALSE)
    ks_stat <- ks.test(ar_data$interarrival_time, "pexp", rate = lambda)$statistic
    ks_pval <- results[1]
    ad_results <- ad.test(ar_data$interarrival_time, "pexp", rate = lambda)
    ad_stat <- ad_results$statistic
    ad_pval <- results[2]
    
    if (ks_pval < 0.05){
      ks_exp_true <- FALSE
    }else {
      ks_exp_true <- TRUE
    }
    
    if (ad_pval < 0.05){
      ad_exp_true <- FALSE
    }else {
      ad_exp_true <- TRUE
    }
  
    results <- c(ar.num, n, lambda, ks_exp_true, ks_stat, ks_pval, ad_exp_true, ad_stat, ad_pval)
    df.30[nrow(df.30) + 1, ] <- results
  }
}

df.30
mean(df.30$ad_exp_true)
mean(df.30$ks_exp_true)
nrow(df.30)
```

```{r, message = FALSE, warning = FALSE}
df.50 <- compData <- data.frame(ar_num = numeric(0), n = numeric(0), lambda = numeric(0), ks_exp_true = numeric(0), ks_stat = numeric(0), ks_pval = numeric(0), ad_exp_true = numeric(0), ad_stat = numeric(0), ad_pval = numeric(0))
for (i in 1:length(unique(GOES_inters$Garreg))){
  
  ar.num <- unique(GOES_inters$Garreg)[i]
  
  ar_data <- GOES_inters %>% filter(Garreg == ar.num)
  # ar_data$interarrival_time <- ar_data$interarrival_time + rnorm(1, 0, 0.0001)
  n <- nrow(ar_data)
  if (n >= 50){
    
    lambda <- 1/mean(ar_data$interarrival_time)
    results <- gtest.sim(ar_data$interarrival_time, n = 10000, boot = FALSE)
    ks_stat <- ks.test(ar_data$interarrival_time, "pexp", rate = lambda)$statistic
    ks_pval <- results[1]
    ad_results <- ad.test(ar_data$interarrival_time, "pexp", rate = lambda)
    ad_stat <- ad_results$statistic
    ad_pval <- results[2]
    
    # try for 0.01
    if (ks_pval < 0.05){
      ks_exp_true <- FALSE
    }else {
      ks_exp_true <- TRUE
    }
    
    if (ad_pval < 0.05){
      ad_exp_true <- FALSE
    }else {
      ad_exp_true <- TRUE
    }
  
    results <- c(ar.num, n, lambda, ks_exp_true, ks_stat, ks_pval, ad_exp_true, ad_stat, ad_pval)
    df.50[nrow(df.50) + 1, ] <- results
  }
}

df.50
mean(df.50$ad_exp_true)
mean(df.50$ks_exp_true)
nrow(df.50)

df.50$ad_exp_true2 <- df.50$ad_pval < 0.01
```



distribution of lambda parameters within ARs
```{r}
GOES_inters_lambda <- GOES_inters %>% group_by(Garreg) %>% summarise(lambda = 1/mean(interarrival_time))

ggplot(data = GOES_inters_lambda, aes(x = lambda)) + 
  geom_histogram(fill = "aquamarine4", color = "black", binwidth = 0.05, boundary = 0) + 
  theme_bw() + 
  labs(x = TeX("$\\lambda$ (Flares/Hour)"), y = "Count")
```

```{r}
ar_nums <- df.50$ar_num
ar_test <- df.50[c(1,7)] %>% filter(ar_num %in% ar_nums)
ar_test_true <- ar_test %>% filter(ad_exp_true2 == 1)
ar_test_true <- ar_test_true[,1]
ar_info <- GOES_inters %>% filter(Garreg %in% ar_nums) %>% mutate(exp_true = ifelse(Garreg %in% ar_test_true2, 1, 0))

ggplot(ar_info, aes(x = interarrival_time, fill = as.factor(exp_true))) + 
  facet_wrap(vars(Garreg), ncol = 5) +
  geom_histogram(color = "black", binwidth = 3, boundary = 0) + 
  theme_bw() + 
  labs(x = TeX("Interarrival Time"), y = "Count")
```

```{r}
ar_info_lambda <- ar_info %>% group_by(Garreg) %>% summarise(start = min(Gstart), end = max(Gstop), lat = mean(Glat[Glat != 0]), n = n(), lambda = 1/mean(interarrival_time), te = mean(Gflrtotalenergy), exp_true = mean(exp_true))

ggplot(ar_info_lambda, aes(x = start, y = n, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Active Region Start Time"), y = "Number of Flares in Active Region", color = "Exponential") +
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = lambda, y = n, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("$\\lambda$ (Flares/Hour)"), y = "Number of Flares in Active Region", color = "Exponential")+
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = start, y = lat, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Active Region Start Time"), y = "Latitude", color = "Exponential")+
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = n, y = lat, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Number of Flares in Active Region"), y = "Latitude", color = "Exponential")+
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = te, y = n, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Average Total Energy"), y = "Number of Flares in Active Region", color = "Exponential") +
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True")) + scale_x_log10()
```

investigation into active region rejections
```{r}
ar_nums <- df.50$ar_num
ar_test <- df.50[c(1,7)] %>% filter(ar_num %in% ar_nums)
ar_test_true <- ar_test %>% filter(ad_exp_true == 1)
ar_test_true <- ar_test_true[,1]
ar_info <- GOES_inters %>% filter(Garreg %in% ar_nums) %>% mutate(exp_true = ifelse(Garreg %in% ar_test_true, 1, 0))

ggplot(ar_info, aes(x = interarrival_time, fill = as.factor(exp_true))) + 
  facet_wrap(vars(Garreg), ncol = 5) +
  geom_histogram(color = "black", binwidth = 3, boundary = 0) + 
  theme_bw() + 
  labs(x = TeX("Interarrival Time"), y = "Count")
```

```{r}
ar_info_lambda <- ar_info %>% group_by(Garreg) %>% summarise(start = min(Gstart), end = max(Gstop), lat = mean(Glat[Glat != 0]), n = n(), lambda = 1/mean(interarrival_time), te = mean(Gflrtotalenergy), exp_true = mean(exp_true))

ggplot(ar_info_lambda, aes(x = start, y = n, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Active Region Start Time"), y = "Number of Flares in Active Region", color = "Exponential") +
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = lambda, y = n, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("$\\lambda$ (Flares/Hour)"), y = "Number of Flares in Active Region", color = "Exponential")+
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = start, y = lat, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Active Region Start Time"), y = "Latitude", color = "Exponential")+
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = n, y = lat, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Number of Flares in Active Region"), y = "Latitude", color = "Exponential")+
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))

ggplot(ar_info_lambda, aes(x = te, y = n, color = as.factor(exp_true))) + 
  geom_point() +
  theme_bw() + 
  labs(x = TeX("Average Total Energy"), y = "Number of Flares in Active Region", color = "Exponential") +
  scale_colour_manual(values = c("red", "green"), breaks = c("0", "1"), labels = c("False", "True"))
```


# 3.5 Discussion

legnth of active region plots with color for lambda, y value for average lattitude with overlaps -> overdispersion from sum of random number of poissons
```{r}
GOES_seg_23 <- GOES %>% filter(cycle == 23 & year > 1996) %>% select(Garreg, year, Gpeak, Gstart, Gstop, Glat) %>% group_by(Garreg) %>% mutate(waiting_time = as.numeric(Gstart-lag(Gstart))) %>% drop_na() %>% summarize(start = min(Gstart), end = max(Gstop), lat = mean(Glat[Glat != 0]), lambda = 1/mean(waiting_time)) 

ggplot(data = GOES_seg_23) + 
  geom_segment(aes(x = start, xend = end, y = lat, yend = lat, color = log(lambda, 10)), size = 4) +
  theme_bw() +
  labs(x = "Date", y = "Latitude", color = TeX("$\\lambda$")) + 
  scale_color_distiller(palette = "Spectral")

GOES_seg_24 <- GOES %>% filter( year == 2013) %>% select(Garreg, year, Gpeak, Gstart, Gstop, Glat) %>% group_by(Garreg) %>% mutate(waiting_time = as.numeric(Gstart-lag(Gstart))) %>% drop_na() %>% summarize(start = min(Gstart), end = max(Gstop), lat = mean(Glat[Glat != 0]), lambda = 1/mean(waiting_time)) 

ggplot(data = GOES_seg_24) + 
  geom_segment(aes(x = start, xend = end, y = lat, yend = lat, color = log(lambda, 10)), size = 1.5) +
  theme_bw() +
  labs(title = TeX("Length of Active Regions and Their $\\hat{\\lambda}$ in 2013"), x = "Date", y = "Latitude", color = TeX("$log(\\lambda)$")) + 
  scale_color_distiller(palette = "Greens")
```


distribution of overall waiting time (thought to follow power law in the tail) (make sure to find waiting times between flares within ARs)
```{r}
ggplot(GOES_inters, aes(x = interarrival_time)) + 
  geom_histogram(fill = "aquamarine4", color = "black", binwidth = 3, boundary = 0) + 
  theme_bw() + 
  labs(x = TeX("Waiting Time"), y = "Count")

ggplot(GOES %>% filter(interarrival_time < 10), aes(x = interarrival_time)) + 
  geom_histogram(fill = "aquamarine4", color = "black", binwidth = 0.2, boundary = 0) + 
  theme_bw() + 
  labs(x = TeX("Waiting Time"), y = "Count")

nrow(GOES %>% filter(interarrival_time <= 1))/nrow(GOES)
```


could explain this overdispersion and aggregate waiting time distribution

## By-Bin Fits

```{r}
# creating a new variable for the observed flares energy_bin
GOES_bin <- GOES %>% filter(year > 1996) %>% select(Gflrtotalenergy, year, cycle, Gstart)

# filtered out 16 flares that occurred on February 29 (leap year days)
GOES_daily <- GOES_bin %>% 
  mutate(day = day(Gstart), month = month(Gstart)) %>% filter(!(day == 29 & month == 2)) %>% 
  select(-day, -month)

# created variable day that gives the calendar day of each flare
GOES_daily <- GOES_daily %>% 
  mutate(day = yday(Gstart)) %>%
  select(-Gstart)

leap_years <- seq(1996, 2020, 4)

break_levels <- levels(cut(seq(1,370,10), breaks = seq(0, 370, 10)))

# for years in leap_years, for days >= 60, subtract 1
# to look at by year counts, do group_by(year)
GOES_daily <- GOES_daily %>% 
  mutate(day = if_else(year %in% leap_years & day >=60, day - 1, day)) %>% 
  mutate(day_bin = cut(day, breaks = seq(0, 370, 10))) %>% mutate(day_bin = factor(day_bin, levels = break_levels)) %>% complete(day_bin)

GOES_year_count <- GOES_daily %>% group_by(year) %>% count(day_bin, name = "n", .drop = F)
```

```{r}
# finding mean and variance for each bin range 1999 to 2002
means <- GOES_year_count %>% group_by(year) %>% summarise(mean = mean(n),var = var(n), lower = qpois(.025, lambda = mean), upper = qpois(.975, lambda = mean))

ggplot(data = GOES_year_count, aes(x = day_bin, y = n)) + 
  geom_point( color = "sienna2") + 
  guides(color = FALSE) +
  facet_wrap(vars(year)) + 
  theme_bw() + 
  labs(title = "10-Day Counts for Each Year with Poisson Fits", subtitle = "Solid line is the mean, horizong lines are 2.5 and 97.5 percentiles of fitted Poisson distribution", x = "Day Bin", y = "Count", color = "Energy Bin") +
  geom_hline(data = means, aes(yintercept = mean)) +
  geom_hline(data = means, aes(yintercept = lower), linetype = "dashed") +
  geom_hline(data = means, aes(yintercept = upper), linetype = "dashed") +
  geom_rug(sides = "l") +
  theme(axis.text.x = element_blank())
```

```{r}
ggplot(GOES_year_count, aes(sample = n)) + 
  facet_wrap(vars(year), scales = "free") +
  stat_pp_point(distribution = "pois",
                color = "aquamarine4") +
  stat_pp_line(color = "sienna2") +
  stat_pp_band(distribution = "pois") +
  theme_bw() +
  labs(title = "P-P Plots for Poisson Fits to Yearly 10-Day Count Distributions", x = "Theoretical Cumulative Probability" , 
       y = "Empirical Cumulative Probability")
```

```{r}
# Simulation-based KS test (for exponential only)... Also known as the Lilliefors Test
ptest.sim <- function(data, n = 10000){
  
  # lf - lilliefors, lf_boot - lilliefors double bootstrap, ad_sim - Anderson-Darling like lilliefors, ad_boot - ad double boot
  
  # mle parameter estimate
  lambda <- mean(data)
  
  # ks p-value from original data
  ks_result <- ks.test(data,"ppois",lambda)$p.value
  ad_result <- ad.test(data, "ppois", lambda)$p.value
    
  # simule data from ~exp(lambda) n times, then find p-value
  sim_values <- replicate(n,
                          {
                            sim_data <- rpois(length(data), lambda = lambda)
                            p_val_ks <- ks.test(sim_data, "ppois", lambda = mean(sim_data))$p.value
                            p_val_ad <- ad.test(sim_data, "ppois", lambda = mean(sim_data))$p.value
                            c(p_val_ks, p_val_ad)
                          })
    
  
  sim_values_ks <- c(sim_values[1,], ks_result)
  sim_values_ad <- c(sim_values[2,], ad_result)
  
  # proportion of p-values <= original ks pval... corrected p-value to report
  sim_pval_ks <- mean(sim_values_ks < ks_result)
  sim_pval_ad <- mean(sim_values_ad < ad_result)
  
  
  return(c(sim_pval_ks, sim_pval_ad))
  
}
```


```{r message=FALSE, warning=FALSE}
means <- GOES_year_count %>% group_by(year) %>% summarise(mean = mean(n),var = var(n), inflation_factor = var(n)/mean(n), lower = qpois(.025, lambda = mean), upper = qpois(.975, lambda = mean))

df_results <- data.frame(year = integer(0), ks_pval = integer(0), ks_reject = integer(0), ad_pval = integer(0), ad_reject = integer(0))

# conducting Lilliefors and AD-Lilliefors test for each year
for (i in 1:nrow(means)){
  
  year <- means[i,]$year
  data <- GOES_year_count[which(GOES_year_count$year == year),]$n
  
  result <- ptest.sim(data, n = 10000)
  # Lilliefors version of KS test
  ks_result <- result[1]
  # A Lilliefors-like method for the Anderson-Darling Test 
  ad_result <- result[2]
  
  reject_ks <- ks_result < 0.05
  reject_ad <- ad_result < 0.05
  
  df_results[i,] <- c(year, ks_result, reject_ks, ad_result, reject_ad)
  
}

df_results
mean(df_results$ks_reject)
mean(df_results$ad_reject)
```

